The YUIMA Project: a Computational Framework for Simulation and Inference of Stochastic Differential Equations

### 2011-09-27 pkg/yuimadocs/inst/doc/JSS/article.Rnw
update
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/yuima@189 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

In the \pkg{yuima} package stochastic differential equations can be of very abstract type, multidimensional, driven by Wiener process of fractional Brownian motion with general Hurst parameter, with or without jumps specified as L\'evy noise.

@@

Stochastic differential equations are 
commonly used 
to model random evolution along continuous or 
practically continuous time, such as 
the random movements of a stock price. 
Theory of statistical inference for 
stochastic differential equations already has a fairly long history, 
more than three decades, but it is still developing quickly new 
methodologies and expanding the area. 
The formulas produced by the theory are usually very sophisticated, 
which makes it difficult for standard users not necessarily 
familiar with this field to enjoy utilities. 
For example, the asymptotic expansion method for computing 
option prices (i.e., expectation of an irregular functional of 
a stochastic process) provides precise approximation values 
instantaneously, taking advantage of the analytic approach, 
but the formula has a long expression like more than one page!

@@

The \code{yuima.model} class contains informations about the stochastic differential equation of interest. The constructor function \code{setModel} is
used to give a mathematical description of the stochastic differential equation. 
All functions in the package are assumed to get as much information
as possible from the model instead of replicating the same code everywhere.
If there are missing pieces of information, we may change or extend the description
of the model.
\\
\\
An object of class \code{yuima.model} contains several slots listed below. 
To see inside its structure, one can use the \proglang{R} command \code{str} on a \pkg{yuima} object.
\begin{itemize}
\item \code{drift} is an \proglang{R} vector of expressions which contains the drift specification;
\item \code{diffusion} is itself a list of 1 slot which describes the diffusion
  coefficient matrix relative to first noise;
\item \code{hurst} is the Hurst index of the fractional Brownian motion, by default \code{0.5}  meaning a standard Brownian motion;
\item \code{parameter} which is a short name for ``parameters'' which 
  is a list with the following entries:
\begin{itemize}
\item \code{all} contains the names of all the parameters found 
   in the diffusion and drift coefficient;
\item \code{common} contains the names of the parameters in common between the drift and diffusion coefficients;
\item \code{diffusion} contains the parameters belonging to the diffusion coefficient;
\item \code{drift} contains the parameters belonging to the drift coefficient;
\item \code{jump} contains the parameters belonging to the  coefficient of the L\'evy noise;
\item \code{measure} contains the parameters describing to the L\'evy measure;
\end{itemize}
\item \code{measure} specifies the measure of the L\'evy noise;
\item \code{measure.type} a switch to specify how the L\'evy measure is described;
\item \code{state.variable} and \code{time.variable}, by default,
  are assumed to be \code{x} and \code{t} but the user can freely choose them.
  The \code{yuima.model} class assumes that the user either use default
  names for \code{state.variable} and \code{time.variable} variables or specify his own names. All
  the rest of the symbols are considered parameters and distributed accordingly
  in the \code{parameter} slot.
\item \code{jump.variable} the name of the variable used in the description of the L\'eve component;
\item \code{solve.variable} contains a vector of variable names, each element corresponds to the
   name of the solution variable (left-hand-side) of each equation in the model, in the corresponding order.
\item \code{noise.number} indicates the number of sources of noise.
\item \code{xinit} initial value of the stochastic differential equation;
\item \code{equation.number} represents the number of equations, i.e. the number of one dimensional
  stochastic differential equations.
\item \code{dimension} reports the dimensions of the parameter space. It is a list of the
  same length of \code{parameter} with the same names.
\item \code{J.flag} for internal use only.
\end{itemize}
As seen in the above, the parameters space is accurately described internally in a \code{yuima} object because in inference for stochastic differential equations, estimators of different parameters have different properties. Usually, the rate of convergence for estimators in the diffusion coefficient are similar to the ones in the i.i.d. sampling while estimators of parameters in the drift coefficient are slower or, in some cases, not even consistent. The \pkg{yuima} always tries to implement the best statistical inference for the given model under the observed sampling scheme.

@@

\section{Asymptotic expansion}\label{sec4}
The \pkg{yuima} package can handle asymptotic expansion of functionals of  $d$-dimensional diffusion process 
$$\de X_t^\ve = a(X_t^\ve,\ve)\de t + b(X_t^\ve,\ve)\de W_t, \qquad \ve \in(0,1]$$
with $W_t$ and $r$-dimensional Wiener process, i.e. $W_t=(W_t^1, \ldots, W_t^r)$.
The functional is expressed in the following abstract form
$$
F^\ve(X_t^\ve)  = \sum_{\alpha=0}^r \int_0^T f_\alpha(X_t^\ve,\de)\de W_t^\alpha + F(X_t^\ve,\ve), \qquad W_t^0=t
$$ 
A typical example of application is the case of Asian option pricing. For example, in the Black \& Scholes model
$$\de X_t^\ve = \mu X_t^\ve\de t + \ve X_t^\ve \de W_t$$ 
the price of the option is of the form
$$ \dE\left\{ \max\left( \frac{1}{T}\int_0^T X_t^\ve \de t - K,0\right)\right\}.$$
Thus the functional of interest is
$$F^\ve(X_t^\ve) = \frac{1}{T}\int_0^T X_t^\ve \de t, \qquad r=1$$ 
with
$$
f_0(x,\ve) = \frac{x}{T}, \quad f_1(x,\ve)=0, \quad  F(x, \ve) = 0
$$
in 
$$
F^\ve(X_t^\ve)  = \sum_{\alpha=0}^r \int_0^T f_\alpha(X_t^\ve,\de)\de W_t^\alpha + F(X_t^\ve,\ve)
$$ 
So, the call option price requires the composition  of a smooth functional
$$F^\ve(X_t^\ve) = \frac{1}{T}\int_0^T X_t^\ve \de t, \qquad r=1$$
with the irregular function
$$
\max(x-K,0)
$$
Monte Carlo methods require a huge number of simulations to get the desired accuracy of the
calculation of the price, while asymptotic expansion of $F^\ve$ provides very accurate approximations.
The \pkg{yuima} package provides functions to construct the functional $F^\ve$, and automatic asymptotic expansion based on Malliavin calculus starting from a \pkg{yuima} object.
Next is an example
<<echo=TRUE, print=FALSE, results=hide>>=
diff.matrix <- matrix( c("x*e"), 1,1)
model <- setModel(drift = c("x"), diffusion = diff.matrix)
T <- 1
xinit <- 1
K <- 1
f <- list( expression(x/T), expression(0))
F <- 0
e <- .3
yuima <- setYuima(model = model, sampling = setSampling(Terminal=T, n=1000))
yuima <- setFunctional( yuima, f=f,F=F, xinit=xinit,e=e)
@
this time the \code{setFunctional} command fills the appropriate slots
<<echo=TRUE, print=FALSE>>=
str(yuima@functional)
@
Then, it is as easy as
<<echo=TRUE, print=FALSE>>=
F0 <- F0(yuima)
F0
max(F0-K,0)  # asian call option price
@
to obtain the zero order approximation of the value of the functional.
We can go up to the first order approximation adding one term to the expansion
<<echo=TRUE, print=FALSE,results=hide>>=
rho <- expression(0)
get_ge <- function(x,epsilon,K,F0){
  tmp <- (F0 - K) + (epsilon * x) 
  tmp[(epsilon * x) < (K-F0)] <- 0
  return( tmp )
}
epsilon <- e  # noise level
g <- function(x) {
  tmp <- (F0 - K) + (epsilon * x) 
 tmp[(epsilon * x) < (K-F0)] <- 0
 tmp
}
asymp <- asymptotic_term(yuima, block=10, rho, g)
@
and the final value is
<<echo=TRUE, print=FALSE>>=
asymp$d0 + e * asymp$d1  # asymp. exp. of asian call price
@

\section{Inference for stochastic processes}\label{sec5}
The \pkg{yuima} implements several optimal techniques for parametric and nonparametric estimation of multidimensional stochastic differential equations.
Although most of the examples in this section are given on simulated data, the main way to fill up the \code{data} slot of a \code{yuima} object is to use the function \code{setYuima}. The function \code{setYuima} sets various slots of the \code{yuima} object. In particular, to estimate a \code{yuima.model} called \code{mod} on the data \code{X} one can use a code like this \code{my.yuima <- setYuima(data=setData(X), model=mod)} and then pass \code{my.yuima} to the inference functions as described in what follows.

\subsection{Quasi Maximum Likelihood estimation}
Consider the multidimensional diffusion process
$$
\de X_t = b(\theta_2,X_t)\de t + \sigma(\theta_1, X_t) \de W_t
$$
where $W_t$ is an {$r$}-dimensional standard Wiener process 
independent of the initial value $X_0=x_0$. 
Quasi-MLE assumes the following approximation of the true log-likelihood for multidimensional diffusions
{\small
\begin{eqnarray}\label{qlik}
\ell_n({\bf X}_n,\theta)
=-\frac12\sum_{i=1}^n\left\{\log\text{det}(\Sigma_{i-1}(\theta_1))
+\frac{1}{\Delta_n}
\Sigma_{i-1}^{-1}(\theta_1)[\Delta X_i-\Delta_n b_{i-1}(\theta_2)]^{\otimes 2}\right\}
\end{eqnarray}}
where $\theta=(\theta_1, \theta_2)$, $\Delta X_i=X_{t_i}-X_{t_{i-1}}$, $\Sigma_i(\theta_1)=\Sigma(\theta_1,X_{t_i})$, $b_i(\theta_2)=b(\theta_2,X_{t_i})$, $\Sigma=\sigma^{\otimes 2}$, $A^{\otimes 2}= A^T A$ and $A^{-1}$ the inverse of $A$, $A[B]^{\otimes 2} = B^T A B$.  Then, \cite{Yoshida92}, the QML estimator of $\theta$ is
$$\tilde\theta_n=\arg\min_\theta \ell_n({\bf X}_n,\theta)$$
As an example, we consider the simple model
\begin{equation}
\de X_t = (2-\theta_2  X_t)  \de t + (1+X_t^2)^{\theta_1}  \de W_t
\label{eq:model1}
\end{equation}
with $\theta_1=0.3$ and $\theta_2=0.2$
<<echo=TRUE, print=FALSE,results=hide>>=
ymodel <- setModel(drift="(2-theta2*x)", diffusion="(1+x^2)^theta1")
n <- 500
ysamp <- setSampling(Terminal = n^(1/3), n = n)
yuima <- setYuima(model = ymodel, sampling = ysamp)
set.seed(123)
yuima <- simulate(yuima, xinit = 1, true.parameter = list(theta1 = 0.2, theta2 = 0.3))
@
With the simulated path we can use the function \code{qmle} to estimate the parameters as follows
<<echo=TRUE, print=FALSE,results=hide>>=
param.init <- list(theta2=0.5,theta1=0.5)
mle1 <- qmle(yuima, start =param.init , 
     lower = list(theta1=0, theta2=0), 
     upper = list(theta1=1, theta2=1))
@
and the estimated coefficients are as follows
<<echo=TRUE, print=FALSE>>=
summary(mle1)
@
Notice the interface and the output of the \code{qmle} is quite similar to the ones of the standard \code{mle} function of the \pkg{stats4} package of the base \proglang{R} system.

@@

Adaptive Bayes estimation is developed in \pkg{yuima} via the method $\tt adaBayes$.
Consider again the model \eqref{eq:model1} with the same values for the parameters, i.e.
 $\theta_1=0.3$ and $\theta_2=0.1$
In order to perform Bayesian estimation, we need to prepare the prior densities for the parameters.
For simplicity we use uniform distributions in $[0,1]$
<<echo=TRUE>>=
prior <- list(theta2=list(measure.type="code",df="dunif(theta2,0,1)"), theta1=list(measure.type="code",df="dunif(theta1,0,1)"))
@
Then we call $\tt adaBayes$ as follows
<<eval=TRUE,echo=TRUE, results=hide>>=
bayes1 <- adaBayes(yuima, start=param.init, prior=prior, method="nomcmc")
@
and we can compare the adaptive Bayes estimates  with the QMLE estimates
<<eval=TRUE,echo=TRUE>>=
coef(bayes1)
coef(mle1)
@
The argument \code{method="nomcmc"} in \code{adaBayes} performs numerical integration, otherwise MCMC method is used.

@@

@@

\subsection{An example of use}
The \code{lasso} method is implemented in the \pkg{yuima} package.
Let us consider the full CKLS model
$$\de X_t = (\alpha+\beta X_t)\de t + \sigma X_t^\gamma\de W_t$$
and let us try to estimate the parameter on  the U.S. Interest Rates monthly data from 06/1964 to 12/1989. We prepare the data, the model and the constraints for optimization
<<results=hide>>=
library(Ecdat)
data(Irates)
rates <- Irates[,"r1"]
plot(rates)
X <- window(rates, start=1964.471, end=1989.333)
mod <- setModel(drift="alpha+beta*x", diffusion=matrix("sigma*x^gamma",1,1))
yuima <- setYuima(data=setData(X), model=mod)
lambda10 <- list(alpha=10, beta =10, sigma =10, gamma =10)
start <- list(alpha=1, beta =-.1, sigma =.1, gamma =1)
low <- list(alpha=-5, beta =-5, sigma =-5, gamma =-5)
upp <- list(alpha=8, beta =8, sigma =8, gamma =8)
@
and now we apply the \code{lasso} function
<<echo=TRUE,results=hide>>=
lasso10 <- lasso(yuima, lambda10, start=start, lower=low, upper=upp,
   method="L-BFGS-B")
@
From which we see that, instead of the general model
$$\de X_t = (\alpha+\beta X_t)\de t + \sigma X_t^\gamma\de W_t$$
<<>>=
round(lasso10$mle, 2)
round(lasso10$lasso, 2)
@
the LASSO method selects the reduced model
$$\de X_t =  0.6 \de t + 0.12 X_t^{\frac32}\de W_t$$

>>>>

In the \pkg{yuima} package stochastic differential equations can be of very abstract type, multidimensional, driven by Wiener process or fractional Brownian motion with general Hurst parameter, with or without jumps specified as L\'evy noise.

@@

Stochastic differential equations are 
commonly used 
to model random evolution along continuous or 
practically continuous time, such as 
the random movements of a stock price. 
Theory of statistical inference for 
stochastic differential equations already has a fairly long history, 
more than three decades, but it is still developing quickly new 
methodologies and expanding the area. 
The formulas produced by the theory are usually very sophisticated, 
which makes it difficult for standard users not necessarily 
familiar with this field to enjoy utilities. 
For example, the asymptotic expansion method for computing 
option prices (i.e., expectation of an irregular functional of 
a stochastic process) provides precise approximation values 
instantaneously, taking advantage of the analytic approach, 
but the formula has a long expression like more than 900 terms!

@@

The \code{yuima.model} class contains information about the stochastic differential equation of interest. The constructor function \code{setModel} is
used to give a mathematical description of the stochastic differential equation. 
All functions in the package are assumed to get as much information
as possible from the model instead of replicating the same code everywhere.
If there are missing pieces of information, we may change or extend the description
of the model.
\\
\\
An object of class \code{yuima.model} contains several slots listed below. 
To see inside its structure, one can use the \proglang{R} command \code{str} on a \pkg{yuima} object.
\begin{itemize}
\item \code{drift} is an \proglang{R} vector of expressions which contains the drift specification;
\item \code{diffusion} is itself a list of 1 slot which describes the diffusion
  coefficient matrix relative to first noise;
\item \code{hurst} is the Hurst index of the fractional Brownian motion, by default \code{0.5}  meaning a standard Brownian motion;
\item \code{parameter} which is a short name for ``parameters'' which 
  is a list with the following entries:
\begin{itemize}
\item \code{all} contains the names of all the parameters found 
   in the diffusion and drift coefficient;
\item \code{common} contains the names of the parameters in common between the drift and diffusion coefficients;
\item \code{diffusion} contains the parameters belonging to the diffusion coefficient;
\item \code{drift} contains the parameters belonging to the drift coefficient;
\item \code{jump} contains the parameters belonging to the  coefficient of the L\'evy noise;
\item \code{measure} contains the parameters describing to the L\'evy measure;
\end{itemize}
\item \code{measure} specifies the measure of the L\'evy noise;
\item \code{measure.type} a switch to specify how the L\'evy measure is described;
\item \code{state.variable} and \code{time.variable}, by default,
  are assumed to be \code{x} and \code{t} but the user can freely choose them.
  The \code{yuima.model} class assumes that the user either use default
  names for \code{state.variable} and \code{time.variable} variables or specify his own names. All
  the rest of the symbols are considered parameters and distributed accordingly
  in the \code{parameter} slot.
\item \code{jump.variable} the name of the variable used in the description of the L\'evy component;
\item \code{solve.variable} contains a vector of variable names, each element corresponds to the
   name of the solution variable (left-hand-side) of each equation in the model, in the corresponding order.
\item \code{noise.number} indicates the number of sources of noise.
\item \code{xinit} initial value of the stochastic differential equation;
\item \code{equation.number} represents the number of equations, i.e. the number of one dimensional
  stochastic differential equations.
\item \code{dimension} reports the dimensions of the parameter space. It is a list of the
  same length of \code{parameter} with the same names.
\item \code{J.flag} for internal use only.
\end{itemize}
As seen in the above, the parameter space is accurately described internally in a \code{yuima} object because in inference for stochastic differential equations, estimators of different parameters have different properties. Usually, the rate of convergence for estimators in the diffusion coefficient are similar to the ones in the i.i.d. sampling while estimators of parameters in the drift coefficient are slower or, in some cases, not even consistent. The \pkg{yuima} always tries to implement the best statistical inference for the given model under the observed sampling scheme.

@@

\section{Asymptotic expansion}\label{sec4}
The \pkg{yuima} package can handle asymptotic expansion of functionals of  $d$-dimensional diffusion process 
$$\de X_t^\ve = a(X_t^\ve,\ve)\de t + b(X_t^\ve,\ve)\de W_t, \qquad \ve \in(0,1]$$
with $W_t$ and $r$-dimensional Wiener process, i.e. $W_t=(W_t^1, \ldots, W_t^r)$.
The functional is expressed in the following abstract form
\begin{equation}
\label{yuima:func1}
F^\ve(X_t^\ve)  = \sum_{\alpha=0}^r \int_0^T f_\alpha(X_t^\ve,\de)\de W_t^\alpha + F(X_t^\ve,\ve), \qquad W_t^0=t.
\end{equation} 
A typical example of application is the case of Asian option pricing. For example, in the Black \& Scholes model
\begin{equation}
\label{eqyuima:gbm1}
\de X_t^\ve = \mu X_t^\ve\de t + \ve X_t^\ve \de W_t
\end{equation}
the price of the option is of the form
$$ \dE\left\{ \max\left( \frac{1}{T}\int_0^T X_t^\ve \de t - K,0\right)\right\}.$$
Thus the functional of interest is
$$F^\ve(X_t^\ve) = \frac{1}{T}\int_0^T X_t^\ve \de t, \qquad r=1$$ 
with
$$
f_0(x,\ve) = \frac{x}{T}, \quad f_1(x,\ve)=0, \quad  F(x, \ve) = 0
$$
in \eqref{yuima:func1}.
So, the call option price requires the composition  of a smooth functional
$$F^\ve(X_t^\ve) = \frac{1}{T}\int_0^T X_t^\ve \de t, \qquad r=1$$
with the irregular function
$$
\max(x-K,0)
$$
Monte Carlo methods require a huge number of simulations to get the desired accuracy of the
calculation of the price, while asymptotic expansion of $F^\ve$ provides very accurate approximations.
The \pkg{yuima} package provides functions to construct the functional $F^\ve$, and automatic asymptotic expansion based on Malliavin calculus \citep{Yoshida92b} starting from a \pkg{yuima} object.
Here is an example. Consider a simple geometric Brownian motion of equation \eqref{eqyuima:gbm1} with $\mu=1$ and $X_0=1$. We define the model and the functional below:
<<echo=TRUE, print=FALSE, results=hide>>=
model <- setModel(drift = "x", diffusion = matrix( "x*e", 1,1))
T <- 1
xinit <- 1
K <- 1
f <- list( expression(x/T), expression(0))
F <- 0
e <- 0.5
yuima <- setYuima(model = model, sampling = setSampling(Terminal=T, n=1000))
yuima <- setFunctional( yuima, f=f,F=F, xinit=xinit,e=e)
@
this time the \code{setFunctional} command fills the appropriate slots inside the \code{yuima} object
<<echo=TRUE, print=FALSE>>=
str(yuima@functional)
@
Then, to obtain the first term in the asymptotic expansion (i.e. for $\ve=0$), it is as easy as calling the function \code{F0} on the \code{yuima} object:
<<echo=TRUE, print=FALSE>>=
F0 <- F0(yuima)
F0
@
so the option price approximation is
<<>>=
max(F0-K,0)  # asian call option price
@
We can go up to the first order approximation adding one term to the expansion
<<echo=TRUE, print=FALSE,results=hide>>=
rho <- expression(0)
epsilon <- e  # noise level
g <- function(x) {
  tmp <- (F0 - K) + (epsilon * x) 
 tmp[(epsilon * x) < (K-F0)] <- 0
 tmp
}
asymp <- asymptotic_term(yuima, block=10, rho, g)
@
and the final value is
<<echo=TRUE, print=FALSE>>=
asymp$d0 + e * asymp$d1  # asymp. exp. of asian call price
@

\section{Inference for stochastic processes}\label{sec5}
The \pkg{yuima} implements several optimal techniques for parametric, semi- and non-parametric estimation of (multidimensional) stochastic differential equations.
Although most of the examples in this section are given on simulated data, the main way to fill up the \code{data} slot of a \code{yuima} object is to use the function \code{setYuima}. The function \code{setYuima} sets various slots of the \code{yuima} object. In particular, to estimate a \code{yuima.model} called \code{mod} on the data \code{X} one can use a code like this \code{my.yuima <- setYuima(data=setData(X), model=mod)} and then pass \code{my.yuima} to the inference functions as described in what follows.

\subsection{Quasi Maximum Likelihood estimation}
Consider the multidimensional diffusion process
$$
\de X_t = b(\theta_2,X_t)\de t + \sigma(\theta_1, X_t) \de W_t
$$
where $W_t$ is an {$r$}-dimensional standard Wiener process 
independent of the initial value $X_0=x_0$. 
Quasi-MLE assumes the following approximation of the true log-likelihood for multidimensional diffusions
{\small
\begin{eqnarray}\label{qlik}
\ell_n({\bf X}_n,\theta)
=-\frac12\sum_{i=1}^n\left\{\log\text{det}(\Sigma_{i-1}(\theta_1))
+\frac{1}{\Delta_n}
\Sigma_{i-1}^{-1}(\theta_1)[\Delta X_i-\Delta_n b_{i-1}(\theta_2)]^{\otimes 2}\right\}
\end{eqnarray}}
where $\theta=(\theta_1, \theta_2)$, $\Delta X_i=X_{t_i}-X_{t_{i-1}}$, $\Sigma_i(\theta_1)=\Sigma(\theta_1,X_{t_i})$, $b_i(\theta_2)=b(\theta_2,X_{t_i})$, $\Sigma=\sigma^{\otimes 2}$, $A^{\otimes 2}= A^T A$ and $A^{-1}$ the inverse of $A$, $A[B]^{\otimes 2} = B^T A B$.  Then, \citep[see e.g.][]{Yoshida92, Kessler97}, the QML estimator of $\theta$ is
$$\tilde\theta_n=\arg\min_\theta \ell_n({\bf X}_n,\theta)$$
As an example, we consider the simple model
\begin{equation}
\de X_t = (2-\theta_2  X_t)  \de t + (1+X_t^2)^{\theta_1}  \de W_t
\label{eq:model1}
\end{equation}
with $\theta_1=0.2$ and $\theta_2=0.3$. We generate sampled data $X_{t_i}$, with $t_i = i \cdot n^{-\frac23}$.
<<echo=TRUE, print=FALSE,results=hide>>=
ymodel <- setModel(drift="(2-theta2*x)", diffusion="(1+x^2)^theta1")
n <- 750
ysamp <- setSampling(Terminal = n^(1/3), n = n)
yuima <- setYuima(model = ymodel, sampling = ysamp)
set.seed(123)
yuima <- simulate(yuima, xinit = 1, true.parameter = list(theta1 = 0.2, theta2 = 0.3))
@
With the sampled data we can use the function \code{qmle} to estimate the parameters as follows
<<echo=TRUE, print=FALSE,results=hide>>=
param.init <- list(theta2=0.5,theta1=0.5)
mle1 <- qmle(yuima, start =param.init , 
     lower = list(theta1=0, theta2=0), 
     upper = list(theta1=1, theta2=1))
@
and the estimated coefficients are extracted from the output object \code{mle1} as follows
<<echo=TRUE, print=FALSE>>=
summary(mle1)
@
Notice the interface and the output of the \code{qmle} is quite similar to the ones of the standard \code{mle} function of the \pkg{stats4} package of the base \proglang{R} system.

@@

Adaptive Bayes estimation is developed in \pkg{yuima} via the method $\tt adaBayes$.
Consider again the model \eqref{eq:model1} with the same values for the parameters.
In order to perform Bayesian estimation, we need to prepare the prior densities for the parameters.
For simplicity we use uniform distributions in $[0,1]$
<<echo=TRUE>>=
prior <- list(theta2=list(measure.type="code",df="dunif(theta2,0,1)"), theta1=list(measure.type="code",df="dunif(theta1,0,1)"))
@
Then we call $\tt adaBayes$, on the same sample data we used for the \code{qmle} function, as follows
<<eval=TRUE,echo=TRUE, results=hide>>=
bayes1 <- adaBayes(yuima, start=param.init, prior=prior)
@
and we can compare the adaptive Bayes estimates  with the QMLE estimates
<<eval=TRUE,echo=TRUE>>=
coef(bayes1)
coef(mle1)
@
%The argument \code{method="nomcmc"} in \code{adaBayes} performs numerical
%integration, otherwise MCMC method is used.

@@

\subsubsection{Asynchronous estimation for nonlinear systems}
Consider now the two-dimensional system with nonlinear feedback
$$
\begin{aligned}
\de X_t &= Y_t\de t + \sigma_1(t,X_t,Y_t) \de W_t\\
\de Y_t &= -X_t\de t + \rho(t,X_t,Y_t)\sigma_2(t,X_t,Y_t) \de W_t+
\sigma_2(t,X_t,Y_t)\sqrt{1-\rho^2(t,X_t,Y_t)}\de B_t
\end{aligned}
$$
with $ \sigma_1(t,X_t,Y_t)=\sqrt{|X_t|(1+t)}$,
$ \sigma_2(t,X_t,Y_t)=\sqrt{|Y_t|}$,
$\rho(t,X_t,Y_t) = \frac{1}{1+X_t^2}$ and $W_t$, $B_t$, two independent
Brownian motions.
We construct the model and we generate data from it
<<>>=
b1 <- function(x,y) y
b2 <- function(x,y) -x
s1 <- function(t,x,y) sqrt(abs(x)*(1+t))
s2 <- function(t,x,y) sqrt(abs(y))
cor.rho <- function(t,x,y) 1/(1+x^2)
diff.mat <- matrix(c("s1(t,x,y)", "s2(t,x,y) * cor.rho(t,x,y)","", "s2(t,x,y) * sqrt(1-cor.rho(t,x,y)^2)"), 2, 2) 
cor.mod <- setModel(drift = c("b1","b2"), diffusion = diff.mat,solve.variable = c("x", "y"),state.var=c("x","y"))

## Generate a path of the process
set.seed(111) 
Terminal <- 1
n <- 10000
yuima.samp <- setSampling(Terminal = Terminal, n = n) 
yuima <- setYuima(model = cor.mod, sampling = yuima.samp) 
yuima <- simulate(yuima, xinit=c(2,3)) 
@
We apply the same Poisson random sampling so that the object \code{Y} will contain  asynchronous data
<<>>=
p1 <- 0.2
p2 <- 0.3
newsamp <- setSampling(
random=list(rdist=c( function(x) rexp(x, rate=p1*n/Terminal), 
function(x) rexp(x, rate=p1*n/Terminal))) )
Y <- subsampling(yuima, sampling = newsamp)
@
<<cceplot3,fig=TRUE,width=9,height=5>>=
plot(Y,main="asynchronous data (non linear case)")
@

The estimated covariance for the complete trajectory \code{yuima}
is now compared with the one obtained on asyncronous data \code{Y}
<<>>=
cce(yuima)
cce(Y)
@

@@

\subsection{An example of use}
The \code{lasso} method is implemented in the \pkg{yuima} package.
Let us consider the full CKLS model
$$\de X_t = (\alpha+\beta X_t)\de t + \sigma X_t^\gamma\de W_t$$
and let us try to estimate the parameter on  the U.S. Interest Rates monthly data from 06/1964 to 12/1989. We prepare the data, the model and the constraints for optimization
<<results=hide>>=
library(Ecdat)
data(Irates)
rates <- Irates[,"r1"]
plot(rates)
X <- window(rates, start=1964.471, end=1989.333)
mod <- setModel(drift="alpha+beta*x", diffusion=matrix("sigma*x^gamma",1,1))
yuima <- setYuima(data=setData(X), model=mod)
lambda10 <- list(alpha=10, beta =10, sigma =10, gamma =10)
start <- list(alpha=1, beta =-.1, sigma =.1, gamma =1)
low <- list(alpha=-5, beta =-5, sigma =-5, gamma =-5)
upp <- list(alpha=8, beta =8, sigma =8, gamma =8)
@
and now we apply the \code{lasso} function
<<echo=TRUE,results=hide>>=
lasso10 <- lasso(yuima, lambda10, start=start, lower=low, upper=upp,
   method="L-BFGS-B")
@
From which we see that, instead of the general model
$$\de X_t = (\alpha+\beta X_t)\de t + \sigma X_t^\gamma\de W_t$$
<<>>=
round(lasso10$mle, 2)
round(lasso10$lasso, 2)
@
the LASSO method selects the reduced model
$$\de X_t =  0.6 \de t + 0.12 X_t^{\frac32}\de W_t.$$
Notice that this model is not an ergodic one, indicating that the LASSO method shows that the real data are indeed not stationary, but still in the family of CKLS models.
